---
title: "Proposal Master thesis"
subtitle: "Residential mobility and neighbourhood choices"
version: "1"
author: "Laura Datta"
date: "19 februari 2020"
output:
  html_document:
    df_print: paged
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Literature 
*see Annex for summary of articles read so far*

#### Concept *residential mobility* explained by

1. Household characteristics
    + socio economic and demographic factors
    + differences mobility rates at different life stages
    + job changes
1. Housing characteristics
    + adapt houses to demands in terms of space, number of rooms
    + age of houses
1. Neighbourhood characteristics
    + neighbourhood quality (facilities)
    + infrastructure
    + urbanisation
1. Housing market characteristics
    + ownership structure
    + housing policies (e.g. local allocation rules)
    + property values
    
## Contributions

Allocation rules for new constructions or social housing are important instruments of a local government's housing policy. To both meet the demand of the citizens in terms of housing wishes as well as to regulate the housing market they need to understand the residential location choices of citizens. Many existing studies on residential mobility look at whether or when somebody is moving. Those studies that look at location choices, do not consider administrative units as granular as neighbourhoods. As most moves occur at short distances (cf...) it is important to understand the drivers of residential location choice at neighbourhood level.

1. Which factors explain neighbourhood choices?
    +   Multinominal choice model
        + How? 
            + All >100 neighbourhoods *or*
            + identify neighbourhood types *or* 
            + higher levels
                + Within neighbourhood
                + Out of neighbourhood, but within municipality
                + Out of municipality, but within region
                + Out of region
            + Clustering of neighbourhoods
                
1. Segmentation of households
    + Latent class model: Identify latent segments to understand the heterogeneity of drivers for neighbourhood choices

1. Most studies on residential mobility focus on explaining why and when households move. Only few studies (so far I haven't found one) attempt to optimse a model which enables to predict where a household is moving to. This, however, is important for local governments in terms of their allocation rules.
    + Prediction of neighbourhood choice using an ML approach

## Available Data

+ movement data for years 2014 - 2018 (even earlier is possible, 2019 will be added), on household level
+ distinction between moving to, within or out of a location (buurt, wijk, gemeente, regio)
+ household characteristics
    + age, sex, ethnicity of household head
    + number of household members
       + if needed characteristics of other household members can be obtained (e.g. age, gender, etc.)
    + type of household (Single person household, couple, family with children etc.)
+ data on **current** and **previous** housing
    + adres
    + date of move
    + Housing characteristics
        + property values
        + renting price if applicable
        + size (m^2, number of rooms)
        + rent vs buy
        + building year
        + type of house
+ data on neighbourhood levels (available from CBS)
    + population composition
    + Leefbarheid
    + Neighbourhood facilities
    
#### Count of moves on neighbourhood levels, Gemeente Den Bosch, 2014 - 2018
```{r, echo = F}
g_moves1418
```

## Questions
+ Model specifications for different models
+ how to integrate both all neighbourhoods as well as whether within or outside Den Bosch into DV
+ Focus on Gemeente Den Bosch or also other municiaplities within the region Noord Oost Brabant


## Annex: Overview literature (to be extended)
